Changes to police demand in COVID-19 pandemic

### 2022-03-23 docs/index.Rmd
Matt's changes to N8 presentation
Changes the slides for the example used to explain why binary comparisons are bad, as well as the example that introduces ARIMA modelling. Also splits the interview quotes so each quote is on a separate slide, re-arranges the conclusion slides and adds a QR code and bitly link in place of the long URL used previously.

---
title: "Changes to police demand in COVID-19 pandemic"
subtitle: "N8PRP Webinar"
author: "Réka Solymosi, Matt Ashby, Eon Kim, Nadia Kennar, Karen Byrom, Alex McMillan, Phoebe Liebelt"
institute: "University of Manchester, Cheshire Constabulary, UCL"
date: "24/03/2022"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false

@@

all_calls <- read_rds("/Volumes/n8_covid/all_calls_forecast.Rds")
call_forecasts <- read_rds("/Volumes/n8_covid/call_forecasts.Rds")
all_rt_forecast <- read_rds("/Volumes/n8_covid/all_resp_time_forecast.Rds")
by_type_rt_forecast <- read_rds("/Volumes/n8_covid/by_type_rt_forecasts.Rds")
by_grade_rt_forecast <- read_rds("/Volumes/n8_covid/by_grade_rt_forecasts.Rds")
all_calls_attended_forecast <- read_rds("/Volumes/n8_covid/all_attended_forecasts.Rds")
by_type_attended_forecast <- read_rds("/Volumes/n8_covid/by_type_attended_forecasts.Rds")
attended_by_grade_forecast <- read_rds("/Volumes/n8_covid/attended_by_grade_forecasts.Rds")
domestic_by_grade_volume_forecasts <- read_rds("/Volumes/n8_covid/domestic_by_grade_volume_forecasts.Rds")
domestic_by_grade_rt_forecast <- read_rds("/Volumes/n8_covid/domestic_by_grade_rt_forecasts.Rds")
domestic_by_grade_attended_forecasts <- read_rds("/Volumes/n8_covid/domestic_by_grade_attended_forecasts.Rds")
by_grade_tos_forecasts <- read_rds("/Volumes/n8_covid/by_grade_tos_forecasts.Rds")
cheshire_lsoas <- st_read("../cheshire_lsoas.geojson")
lsoa_filler <- bind_rows(data.frame(lsoa = unique(cheshire_lsoas$LSOA11NM),
                          pre_post = "Pre-COVID"), 
                         data.frame(lsoa = unique(cheshire_lsoas$LSOA11NM),
                          pre_post = "2020"))
total_calls_by_imd <- read_csv("/Volumes/n8_covid/total_calls_by_imd.csv")
weekly_count_by_lsoa <- read_csv("/Volumes/n8_covid/weekly_count_by_lsoa.csv")
weekly_asb_by_lsoa <- read_csv("/Volumes/n8_covid/weekly_asb_by_lsoa.csv")
# gini_asb <- read_rds("/Volumes/n8_covid/gini_asb_forecast.Rds")

@@

> Binary comparisons of one period to another are common, but can be seriously misleading.

---
.center[![COVID19 and crime article](imgs/cov_news_01.png)]

---
class: center, middle
.center[![COVID19 and crime article](imgs/cov_news_03.png)]

---
.center[![COVID19 and crime article](imgs/cov_news_02.png)]
---

.center[![COVID19 and crime article](imgs/cov_news_04.png)]

@@

# Can miss the big picture
.center[![miss picture](imgs/miss_whole_picture.png)]

---
# Important to zoom out
.center[![miss picture 2](imgs/miss_pic_2.png)]

---
# Need to account for trends
.center[![trends](imgs/ignore_trends.png)]

@@

---
class: middle
# Forecasting: a better method

.pull-left[Create a synthetic comparison group by using previous crime data to build a model that forecast how much crime would have occurred in the absence of the pandemic.]
.pull-right[![forecasting](imgs/forecasting_book.png)]

@@

# Forecasting Calls for Service

@@

# Quiet roads

```{r quietroadsplots, fig.width=11,fig.height=8}

p1 <- forecast_chart(call_forecasts, "incident_type_new" ,"traffic collision") + 
  labs(title = "Traffic Collision",
      subtitle = "",
      caption = "",
      x = NULL,
      y = NULL,
      fill = NULL) + 
  guides(fill="none")
p2 <- forecast_chart(call_forecasts, "incident_type_new" ,"Highway Disruption")+ 
  labs(subtitle = "", caption = "", x = NULL, fill = NULL) + 
  guides(fill="none")
p3 <- forecast_chart(call_forecasts, "incident_type_new" ,"Road Related Traffic Offence")  +
  labs(subtitle = "",
      caption = "Forecast calculated using data up to 31 January 2020",
      x = NULL,
      y = NULL,
      fill = "Actual observed value significantly different from forecast")

gridExtra::grid.arrange(p1, p2, p3, ncol = 1)

@@

---
# Expressions of concern

> "ASB calls [are] about COVID breaches being coded as ASB... lots of people were reporting other people for breaching the rules. There were a lot of calls where people were very frustrated and they were complaining about their neighbours and the sort of call that sticks in my mind, which I found quite moving quite upsetting really was when somebody would ring in and say that they weren't able to go and visit their elderly relative who lived maybe a mile away... and their next door neighbour was having a party and all these people were turning up." (interview 1)

--

> "... at the start of it, nobody really knew how it was going to affect them. I think there were a lot of people were quite genuinely scared that they were gonna maybe die from it." (interview 3)

--

@@

# Calls for information/ reassurance

> "...in the first part of the first lockdown, people were very confused and very worried about breaking the law and very worried about being arrested for and they would ring up and say, can I do this and can I do that so people were actually asking our advice ..." (interview 1)

--

> "...there are also people calling in because they wanted advice about regulations or were uncertain about what they were allowed to do things..." (interview 3)

--

> "I'd say there's definitely an increase in mental health related incidents that we're seeing and I think people are just really struggling with the isolation of being away from everyone... Sometimes people don't know [who to call] to speak to someone for support so they just ring the police. " (interview 4)

@@

# Presence of guardians/ absence of targets

```{r guardiansplot1, fig.width=11,fig.height=8}

forecast_chart(call_forecasts, "incident_type_new", "Burglary - Residential")

@@

# Presence of guardians/ absence of targets

```{r guardiansplot2, fig.width=11,fig.height=7}

p1 <- forecast_chart(call_forecasts, "incident_type_new", "Theft From Motor Vehicle") + 
  labs(subtitle = "", caption = "", x = NULL, y = NULL, fill = NULL) + 
  guides(fill="none")
p2 <- forecast_chart(call_forecasts, "incident_type_new", "Theft Of Motor Vehicle") + 
  labs(subtitle = "", caption = "", x = NULL, y = NULL, fill = NULL) + 
  guides(fill="none")
p3 <- forecast_chart(call_forecasts, "incident_type_new", "Theft Other")

gridExtra::grid.arrange(p1, p2,p3, ncol = 1)

@@

# Presence of guardians/ absence of targets

```{r guardiansplot3, fig.width=11,fig.height=7}

p1 <- forecast_chart(call_forecasts, "incident_type_new", "Shoplifting") + 
  labs(subtitle = "", caption = "", x = NULL, y = NULL, fill = NULL) + 
  guides(fill="none")
p2 <- forecast_chart(call_forecasts, "incident_type_new", "Robbery")

@@

--

@@

> "There's definitely more online crime. So obviously people are not able to go out and interact in person, but they can over social media and other avenues, such as phone calls, texts - these definitely seem more... I'd say they changed in the sense that the threats were still being made, but not in person, so they will be made over a text or a phone call, social media..." (interview 4)

@@

---
class: center, middle
# Domestic incidents: no increase

@@

--

@@

--

> "... when we get the calls it's like ... when they've been out on the town up to the pubs having drinks and then they come back and they disagree over something so little. And that's when the domestic start." (interview 5)

---
class: center, middle
# Domestic incidents: couldn't call?

> "I remember a lady who was desperately upset. She was very, very distressed and this must have been as lockdown was easing a bit because her husband had actually gone out and this was the first opportunity she'd had to make the call... that Lady had more than likely gone through all sorts before she could actually found the opportunity to ring it in." (interview 1)

---
class: center, middle
# Domestic incidents: change in type

> "I'd say there's quite a lot of child related incidents to deal with domestic ... You know this parents kept this child too long." (interview 4)

--

> "The other thing as well that we found uhm with the kids as well. People keeping their teenagers inside the house. You know, quite a lot of the kids were getting a bit rebellious and they wanted to go out with their mates and (...) they were having arguments with mum and dad about the fact they couldn't leave the house. There was a a few more of those and you know, sort of." (interview 8)

---
class: center, middle, inverse
# Police response

---
class: center, middle
# Attendance rate

```{r allattendedfig, fig.width=11,fig.height=7}

# % attended fig and response time fig
all_calls_attended_forecast %>% 
    mutate(forecast_lower = ifelse(forecast_lower < 0, 0, forecast_lower)) %>% 
    ggplot() +
    # Forecast
    geom_ribbon(
      aes(incident_week, ymin = forecast_lower, ymax = forecast_upper), 
      na.rm = TRUE,
      alpha = 0.5, 
      fill = "grey80"
    ) +
    geom_line(aes(incident_week, forecast_mean), na.rm = TRUE, linetype = "22") +
    # Dates of interest
    geom_vline(aes(xintercept = date), data = dates, linetype = "12") +
    geom_label(aes(date, 0, label = row), data = dates, colour = "grey20") +
    # Actual calls
    geom_line(aes(incident_week, actual_attended)) +
    geom_point(aes(incident_week, actual_attended, fill = sig), shape = 21) +
    scale_x_date(date_labels = "%e %b\n%Y", 
               limits = as.Date(c("2020-01-06", "2020-12-21"))) +
    scale_y_continuous(limits = c(0, NA), labels = scales::comma_format()) +
    scale_fill_manual(values = c(`TRUE` = "black", `FALSE` = "grey80")) +
    labs(
      title = "Attendance rate for all calls for service during 2020 compared to pre-pandemic forecast",
      subtitle = str_wrap(
        str_glue("Events by week: ", str_c(pull(dates, label), collapse = "; ")), 
        80
      ),
      caption = "Forecast calculated using data up to 31 January 2020",
      x = NULL,
      y = "weekly % of calls attended",
      fill = "actual % significantly different from forecast"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom", plot.title = element_text(face = "bold"))

---
class: center, middle
# Attendance rate

```{r bygradeattendedperc, fig.width=11,fig.height=7}

p1 <- forecast_chart(attended_by_grade_forecast,"initial_grade_of_response", "Grade 1") + 
  labs(subtitle = "", caption = "", x = NULL, fill = NULL) + 
  guides(fill="none")
p2 <- forecast_chart(attended_by_grade_forecast,"initial_grade_of_response", "Grade 2") + 
  labs(subtitle = "", caption = "", x = NULL, y = NULL, fill = NULL) + 
  guides(fill="none")
p3 <- forecast_chart(attended_by_grade_forecast,"initial_grade_of_response", "Grade 3") + 
  labs(subtitle = "", caption = "", x = NULL, y = NULL, fill = NULL) + 
  guides(fill="none")
p4 <- forecast_chart(attended_by_grade_forecast,"initial_grade_of_response", "Grade 4") + 
  labs(subtitle = "", caption = "", x = NULL, y = NULL, fill = NULL) + 
  guides(fill="none")

gridExtra::grid.arrange(p1, p2, p3, p4, ncol = 2)

---
class: center, middle
# Response time

```{r allresptimefig, fig.width=11,fig.height=8}

all_rt_forecast %>% 
    mutate(forecast_lower = ifelse(forecast_lower < 0, 0, forecast_lower)) %>% 
    ggplot() +
    # Forecast
    geom_ribbon(
      aes(incident_week, ymin = forecast_lower, ymax = forecast_upper), 
      na.rm = TRUE,
      alpha = 0.5, 
      fill = "grey80"
    ) +
    geom_line(aes(incident_week, forecast_mean), na.rm = TRUE, linetype = "22") +
    # Dates of interest
    geom_vline(aes(xintercept = date), data = dates, linetype = "12") +
    geom_label(aes(date, 0, label = row), data = dates, colour = "grey20") +
    # Actual calls
    geom_line(aes(incident_week, actual_rt)) +
    geom_point(aes(incident_week, actual_rt, fill = sig), shape = 21) +
    scale_x_date(date_labels = "%e %b\n%Y") +
    scale_y_continuous(limits = c(0, NA), labels = scales::comma_format()) +
    scale_fill_manual(values = c(`TRUE` = "black", `FALSE` = "grey80")) +
    labs(
      title = "Median response time for all calls for service during 2020 compared to pre-pandemic forecast",
      caption = "Forecast calculated using data up to 31 January 2020",
      x = NULL,
      y = "Median response time in minutes",
      fill = "Response time significantly different from forecast"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom", plot.title = element_text(face = "bold"))

---
class: center, middle
# Response time

```{r bygraderesptime, fig.width=11,fig.height=7}

p1 <- forecast_chart(by_grade_rt_forecast,"initial_grade_of_response", "Grade 1") + 
  labs(subtitle = "", caption = "", x = NULL, fill = NULL) + 
  guides(fill="none")
p2 <- forecast_chart(by_grade_rt_forecast,"initial_grade_of_response", "Grade 2") + 
  labs(subtitle = "", caption = "", x = NULL, y = NULL, fill = NULL) + 
  guides(fill="none")
p3 <- forecast_chart(by_grade_rt_forecast,"initial_grade_of_response", "Grade 3") + 
  labs(subtitle = "", caption = "", x = NULL, y = NULL, fill = NULL) + 
  guides(fill="none")
p4 <- forecast_chart(by_grade_rt_forecast,"initial_grade_of_response", "Grade 4") + 
  labs(subtitle = "", caption = "", x = NULL, y = NULL, fill = NULL) + 
  guides(fill="none")

gridExtra::grid.arrange(p1, p2, p3, p4, ncol = 2)

---
class: center, middle
# Time on scene

```{r bygradetos, fig.width=11,fig.height=7}

p1 <- forecast_chart(by_grade_tos_forecasts,"initial_grade_of_response", "Grade 1") + 
  labs(subtitle = "", caption = "", x = NULL, fill = NULL) + 
  guides(fill="none")
p2 <- forecast_chart(by_grade_tos_forecasts,"initial_grade_of_response", "Grade 2") + 
  labs(subtitle = "", caption = "", x = NULL, y = NULL, fill = NULL) + 
  guides(fill="none")
p3 <- forecast_chart(by_grade_tos_forecasts,"initial_grade_of_response", "Grade 3") + 
  labs(subtitle = "", caption = "", x = NULL, y = NULL, fill = NULL) + 
  guides(fill="none")
p4 <- forecast_chart(by_grade_tos_forecasts,"initial_grade_of_response", "Grade 4") + 
  labs(subtitle = "", caption = "", x = NULL, y = NULL, fill = NULL) + 
  guides(fill="none")

gridExtra::grid.arrange(p1, p2, p3, p4, ncol = 2)

---
class: center, middle, inverse
# Change in nature of the demand

@@

# Increase of calls recorded as ASB (concern over COVID-19 breach) from less deprived areas

@@

# Overall

>>>>

---
title: "Changes to police demand in COVID-19 pandemic"
subtitle: "N8PRP Webinar"
author: "Réka Solymosi, Matt Ashby, Eon Kim, Nadia Kennar, Karen Byrom, Alex McMillan, Phoebe Liebelt"
institute: "University of Manchester, Cheshire Constabulary, UCL"
date: "24/03/2022"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
class: center, middle

@@

```{r load-data, include=FALSE}
all_calls <- read_rds("/Volumes/n8_covid/all_calls_forecast.Rds")
call_forecasts <- read_rds("/Volumes/n8_covid/call_forecasts.Rds")
all_rt_forecast <- read_rds("/Volumes/n8_covid/all_resp_time_forecast.Rds")
by_type_rt_forecast <- read_rds("/Volumes/n8_covid/by_type_rt_forecasts.Rds")
by_grade_rt_forecast <- read_rds("/Volumes/n8_covid/by_grade_rt_forecasts.Rds")
all_calls_attended_forecast <- read_rds("/Volumes/n8_covid/all_attended_forecasts.Rds")
by_type_attended_forecast <- read_rds("/Volumes/n8_covid/by_type_attended_forecasts.Rds")
attended_by_grade_forecast <- read_rds("/Volumes/n8_covid/attended_by_grade_forecasts.Rds")
domestic_by_grade_volume_forecasts <- read_rds("/Volumes/n8_covid/domestic_by_grade_volume_forecasts.Rds")
domestic_by_grade_rt_forecast <- read_rds("/Volumes/n8_covid/domestic_by_grade_rt_forecasts.Rds")
domestic_by_grade_attended_forecasts <- read_rds("/Volumes/n8_covid/domestic_by_grade_attended_forecasts.Rds")
by_grade_tos_forecasts <- read_rds("/Volumes/n8_covid/by_grade_tos_forecasts.Rds")
cheshire_lsoas <- st_read("../cheshire_lsoas.geojson")
lsoa_filler <- bind_rows(data.frame(lsoa = unique(cheshire_lsoas$LSOA11NM),
                          pre_post = "Pre-COVID"), 
                         data.frame(lsoa = unique(cheshire_lsoas$LSOA11NM),
                          pre_post = "2020"))
total_calls_by_imd <- read_csv("/Volumes/n8_covid/total_calls_by_imd.csv")
weekly_count_by_lsoa <- read_csv("/Volumes/n8_covid/weekly_count_by_lsoa.csv")
weekly_asb_by_lsoa <- read_csv("/Volumes/n8_covid/weekly_asb_by_lsoa.csv")
# gini_asb <- read_rds("/Volumes/n8_covid/gini_asb_forecast.Rds")

@@

Binary comparisons of one period to another are common, but can be seriously misleading.

@@

class: center, middle
.pull-right[
![COVID19 and crime article](imgs/cov_news_01.png)
![COVID19 and crime article](imgs/cov_news_02.png)
]

.pull-left[
![COVID19 and crime article](imgs/cov_news_03.png)
![COVID19 and crime article](imgs/cov_news_04.png)]

# It's easy to miss the big picture

```{r atlanta-map-prep, include=FALSE}
tidy_data <- read_csv("imgs/tracking-crime-changes-data.rds") %>% 
  filter(offense_type == "HOMICIDE") %>% 
  mutate(year = year(offense_date)) %>% 
  count(year, wt = n) %>% 
  filter(year %in% 2016:2017) %>% 
  mutate(
    change = (n - lag(n)) / lag(n),
    hjust = ifelse(year == 2016, 1, 0),
    label = ifelse(
      hjust == 0, 
      str_glue("{n} homicides\n({scales::percent(change)})"), 
      str_glue("{n} homicides")
    )
  )

example_homicides <- read_rds("imgs/atlanta_example_homicides.Rds")
```

```{r atlanta-map-1, fig.asp=0.56, fig.retina=4, out.width="100%"}
# Create an empty map with necessary scales, theme, etc. (geoms will be added
# later)
atlanta_map <- ggplot(tidy_data, aes(year, n)) +
  scale_x_continuous(
    breaks = 2016:2017, 
    limits = c(2010, 2018),
    labels = scales::number_format(accuracy = 1, big.mark = ""),
    expand = expansion(mult = c(0.025, 0.1))
  ) +
  scale_y_continuous(
    limits = c(0, 150), 
    expand = expansion(mult = c(0, 0.03))
  ) +
  labs(
    x = NULL,
    y = "annual count of homicides",
    caption = "Data: Atlanta Police Department"
  ) +
  theme_minimal() +
  theme(panel.grid.minor.x = element_blank())

atlanta_map +
  geom_line(na.rm = TRUE, colour = "#EA7600", size = 1) +
  geom_point(na.rm = TRUE, colour = "#EA7600") +
  geom_label(
    aes(hjust = hjust, label = label), 
    fill = NA, 
    label.padding = unit(0.5, "lines"), 
    label.size = NA,
    lineheight = 1.05
  )
```

# It's easy to miss the big picture

```{r atlanta-map-2, fig.asp=0.56, fig.retina=4, out.width="100%"}
trend_type <- "part of a downward trend"

atlanta_map +
	geom_line(
	  data = filter(example_homicides, scenario == trend_type), 
	  na.rm = TRUE,
	  colour = "grey80", 
	  size = 1
	) +
	geom_point(
	  data = filter(example_homicides, scenario == trend_type), 
	  na.rm = TRUE,
	  colour = "grey80"
	) +
  geom_line(na.rm = TRUE, colour = "#EA7600", size = 1) +
  geom_point(na.rm = TRUE, colour = "#EA7600") +
  annotate(
    "label", 
    label = trend_type,
    x = 2010.5, 
    y = 10, 
    fill = "orange", 
    fontface = "bold",
    hjust = 0, 
    label.size = NA, 
    label.padding = unit(0.5, "lines"), 
    size = 5,
    vjust = 0
  )
```

# It's easy to miss the big picture

```{r atlanta-map-3, fig.asp=0.56, fig.retina=4, out.width="100%"}
trend_type <- "a one-year blip with no trend"

atlanta_map +
	geom_line(
	  data = filter(example_homicides, scenario == trend_type), 
	  na.rm = TRUE,
	  colour = "grey80", 
	  size = 1
	) +
	geom_point(
	  data = filter(example_homicides, scenario == trend_type), 
	  na.rm = TRUE,
	  colour = "grey80"
	) +
  geom_line(na.rm = TRUE, colour = "#EA7600", size = 1) +
  geom_point(na.rm = TRUE, colour = "#EA7600") +
  annotate(
    "label", 
    label = trend_type,
    x = 2010.5, 
    y = 10, 
    fill = "orange", 
    fontface = "bold",
    hjust = 0, 
    label.size = NA, 
    label.padding = unit(0.5, "lines"), 
    size = 5,
    vjust = 0
  )
```

# It's easy to miss the big picture

```{r atlanta-map-4, fig.asp=0.56, fig.retina=4, out.width="100%"}
trend_type <- "a one-year blip in an upward trend"

atlanta_map +
	geom_line(
	  data = filter(example_homicides, scenario == trend_type), 
	  na.rm = TRUE,
	  colour = "grey80", 
	  size = 1
	) +
	geom_point(
	  data = filter(example_homicides, scenario == trend_type), 
	  na.rm = TRUE,
	  colour = "grey80"
	) +
  geom_line(na.rm = TRUE, colour = "#EA7600", size = 1) +
  geom_point(na.rm = TRUE, colour = "#EA7600") +
  annotate(
    "label", 
    label = trend_type,
    x = 2010.5, 
    y = 10, 
    fill = "orange", 
    fontface = "bold",
    hjust = 0, 
    label.size = NA, 
    label.padding = unit(0.5, "lines"), 
    size = 5,
    vjust = 0
  )
```

# It's easy to miss the big picture

```{r atlanta-map-5, fig.asp=0.56, fig.retina=4, out.width="100%"}
trend_type <- "random variation with no trend"

atlanta_map +
	geom_line(
	  data = filter(example_homicides, scenario == trend_type), 
	  na.rm = TRUE,
	  colour = "grey80", 
	  size = 1
	) +
	geom_point(
	  data = filter(example_homicides, scenario == trend_type), 
	  na.rm = TRUE,
	  colour = "grey80"
	) +
  geom_line(na.rm = TRUE, colour = "#EA7600", size = 1) +
  geom_point(na.rm = TRUE, colour = "#EA7600") +
  annotate(
    "label", 
    label = trend_type,
    x = 2010.5, 
    y = 10, 
    fill = "orange", 
    fontface = "bold",
    hjust = 0, 
    label.size = NA, 
    label.padding = unit(0.5, "lines"), 
    size = 5,
    vjust = 0
  )
```

# Need to account for trends
.center[![trends](imgs/ignore_trends.png)]

@@

@@

# Another way to understand change

```{r cin-chart-prep, include=FALSE}
# Load data from Cincinnati for ARIMA example chart
cin_calls <- read_rds("imgs/cin_calls.Rds")
cin_model <- read_rds("imgs/cin_model.Rds")
cin_forecast <- read_rds("imgs/cin_forecast.Rds")

# Build a basic chart that we can build on later
cin_chart <- cin_calls %>% 
  filter(week <= ymd("2020-01-20")) %>% 
  ggplot(aes(week, calls)) +
  scale_x_date(
    date_breaks = "6 months", 
    date_labels = "%b\n%Y", 
    limits = c(ymd("2016-01-01"), ymd("2020-09-30")),
    expand = expansion(mult = c(0.02, 0))
  ) +
  scale_y_continuous(
    limits = c(0, 11000), 
    labels = scales::comma_format(accuracy = 1, scale = 1/1000, suffix = "k"),
    expand = expansion(mult = c(0, 0.05))
  ) +
    scale_fill_manual(
      values = c(`FALSE` = "grey90", `TRUE` = "black"),
      labels = c(
        `FALSE` = "calls within 99% CI",
        `TRUE` = "calls outside 99% CI"
      )
    ) +
  labs(
    fill = NULL,
    x = NULL,
    y = "total calls per week, Cincinnati"
  ) +
  theme_minimal() +
  theme(
    legend.background = element_rect(colour = NA, fill = "white"),
    legend.justification = c(1, 0),
    legend.position = c(1, 0.05),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )
```

```{r cin-chart-1, fig.asp=0.56, fig.retina=4, out.width="100%"}
# Print the basic chart (this needs to be in a seprate chunk because the 
# `cin-chart-prep` chunk needs `include=FALSE`)
cin_chart
```

# Another way to understand change

```{r cin-chart-2, fig.asp=0.56, fig.retina=4, out.width="100%"}
cin_chart <- cin_chart + 
  # print points showing actual counts before COVID
  geom_point(size = 0.75, colour = "grey67")

cin_chart +
  annotate(
    "label", 
    label = "1. count calls",
    x = ymd("2016-03-01"), 
    y = 1000, 
    fill = "orange", 
    fontface = "bold", 
    hjust = 0, 
    label.size = NA, 
    label.padding = unit(0.5, "lines"), 
    vjust = 0
  )
```

# Another way to understand change

```{r cin-chart-3, fig.asp=0.56, fig.retina=4, out.width="100%"}
cin_chart <- cin_chart + 
  # print line showing model fitted counts
  geom_line(aes(week, .fitted), data = cin_model)

cin_chart + 
  annotate(
    "label", 
    label = "2. estimate auto-ARIMA model",
    x = ymd("2016-03-01"), 
    y = 1000, 
    fill = "orange", 
    fontface = "bold", 
    hjust = 0, 
    label.size = NA, 
    label.padding = unit(0.5, "lines"), 
    vjust = 0
  )
```

# Another way to understand change

```{r cin-chart-4, fig.asp=0.56, fig.retina=4, out.width="100%"}
cin_chart <- cin_chart + 
  geom_vline(
    aes(xintercept = ymd("2020-01-20")), 
    size = 0.25, 
    colour = "grey50"
  ) +
  # print line showing CI around estimate
  geom_ribbon(
    aes(x = week, ymin = lower, ymax = upper), 
    data = cin_forecast,
    na.rm = TRUE,
    inherit.aes = FALSE,
    fill = "grey90"
  ) +
  # print line showing estimate of crimes
  geom_line(
    aes(week, estimate), 
    data = cin_forecast, 
    na.rm = TRUE, 
    colour = "grey60",
    linetype = "32"
  )

cin_chart + 
  annotate(
    "label", 
    label = "3. forecast from model (with 99% CIs)",
    x = ymd("2016-03-01"), 
    y = 1000, 
    fill = "orange", 
    fontface = "bold", 
    hjust = 0, 
    label.size = NA, 
    label.padding = unit(0.5, "lines"), 
    vjust = 0
  )
```

# ARIMA models: a better way to understand change

```{r cin-chart-5, fig.asp=0.56, fig.retina=4, out.width="100%"}
cin_chart + 
  # print line showing count of crimes
  geom_line(
    aes(x = week, y = calls), 
    data = filter(cin_forecast, week >= ymd("2020-01-20")), 
    na.rm = TRUE, 
    colour = "black"
  ) +
  # print points showing actual counts after COVID
  geom_point(
    aes(week, calls, fill = sig), 
    data = filter(cin_forecast, week >= ymd("2020-01-20")), 
    na.rm = TRUE,
    # size = 0.75, 
    colour = "black",
    shape = 21
  ) + 
  annotate(
    "label", 
    label = "4. compare counts to forecast",
    x = ymd("2016-03-01"), 
    y = 1000, 
    fill = "orange", 
    fontface = "bold", 
    hjust = 0, 
    label.size = NA, 
    label.padding = unit(0.5, "lines"), 
    vjust = 0
  )
```

@@

@@

> "ASB calls [are] about COVID breaches being coded as ASB … lots of people were reporting other people for breaching the rules. There were a lot of calls where people were very frustrated and they were complaining about their neighbours and the sort of call that sticks in my mind, which I found quite moving quite upsetting really was when somebody would ring in and say that they weren't able to go and visit their elderly relative who lived maybe a mile away … and their next door neighbour was having a party and all these people were turning up." (interview 1)

> "… at the start of it, nobody really knew how it was going to affect them. I think there were a lot of people were quite genuinely scared that they were gonna maybe die from it." (interview 3)

@@

# Calls for information/reassurance

> "… in the first part of the first lockdown, people were very confused and very worried about breaking the law and very worried about being arrested for and they would ring up and say, can I do this and can I do that so people were actually asking our advice …" (interview 1)

@@

# Calls for information/reassurance

> "… there are also people calling in because they wanted advice about regulations or were uncertain about what they were allowed to do things …" (interview 3)

@@

# Calls for information/reassurance

> "I'd say there's definitely an increase in mental health related incidents that we're seeing and I think people are just really struggling with the isolation of being away from everyone … Sometimes people don't know [who to call] to speak to someone for support so they just ring the police. " (interview 4)

@@

# Support for such calls

> "We were directing people to the government website … We were we're not really qualified to advise people on the regulations that actually fit their particular area, because there was an area to area thing at certain times during the lockdown where different local councils and had different idea about the rules, interpreted the rules differently …" (interview 1)

# Support for such calls

> "Getting quite a few calls from elderly people asking for advice … advice on guidelines … Of course we referred them to the government website because they're better off going something like that's got advice in relation to [COVID regulations] (interview 4)

# Support for such calls

> "We have an FAQ. That specially for FCC stuff. And most of the answers can be found in there … There was plenty of support here at work … we had a very large VDU display on the wall with the latest rules and regulations like a like a PowerPoint running through on a continual on a continual rotation." (interview 1)

@@

@@

> "There's definitely more online crime. So obviously people are not able to go out and interact in person, but they can over social media and other avenues, such as phone calls, texts - these definitely seem more … I'd say they changed in the sense that the threats were still being made, but not in person, so they will be made over a text or a phone call, social media …" (interview 4)

@@

# Domestic incidents: no increase

@@

# Domestic incidents: no increase

@@

# Domestic incidents: no increase

> "… when we get the calls it's like … when they've been out on the town up to the pubs having drinks and then they come back and they disagree over something so little. And that's when the domestic start." (interview 5)

# Domestic incidents: couldn't call?

> "I remember a lady who was desperately upset. She was very, very distressed and this must have been as lockdown was easing a bit because her husband had actually gone out and this was the first opportunity she'd had to make the call … that Lady had more than likely gone through all sorts before she could actually found the opportunity to ring it in." (interview 1)

# Change in nature of the demand

@@

# ASB (concern over COVID-19 breach)

@@

---
class: center, middle, inverse
# Conclusions

# Highlights

- Expressions of concern
- Calls for advice
- Violence drop
- No change in domestic violence
- For ASB the caller "profile" changed

---
class: center, middle
# Recommendations

- Further spatio-temporal analysis
- Share good practice (e.g. FHQ information to clarify rules and regulations)
- Further and more robust look at changes to domestic incidents

# More details

Much more (e.g. quiet roads, absence of potential victims, presence of guardians, changes in police response) at [**bit.ly/n8-covid-calls-report**](https://bit.ly/n8-covid-calls-report)

<img src="imgs/n8-covid-calls-report-qr-code.png" alt="QR code leading to the URL https://bit.ly/n8-covid-calls-report", height="300">

---
class: center, middle
# Further training outputs

- [UKDS webinar: an introduction to time series analysis and forecasting](https://ukdataservice.ac.uk/events/an-introduction-to-time-series-analysis-and-forecasting-2/)
- A co-developed series for analysts using R and Python
